Supplementary Materials from: \textit{Innovation and elaboration on the avian tree of life}; Science Advances, in review

### 2023-06-05 Manuscript/main.tex
Update main.tex
PGLS table
Added pgls table for BM results. Formatting needs fixing.

Widely documented, megaevolutionary jumps in phenotypic diversity continue to perplex researchers because it remains unclear whether these dramatic changes can emerge from microevolutionary processes.
Here we tackle this question using new approaches for modeling multivariate traits %@L_multi_traits
 to evaluate the magnitude and distribution of elaboration and innovation in the evolution of bird beaks.
We find that elaboration, evolution along the major axis of phenotypic change, is common at both macro- and megaevolutionary scales whereas innovation, evolution away from the major axis of phenotypic change, is more prominent at megaevolutionary scales.
Indeed, the major axis of phenotypic change among species beak shapes at megaevolutionary scales is an emergent property of innovation across clades. Our analyses suggest that the reorientation of phenotypes via innovation is a ubiquitous route for divergence that can arise through gradual change alone, opening up new avenues for evolution to explore.

@@

\begin{table}
\begin{tabular}{ c c c }
 So & many & stars \\ 
 *** & *** & *** \\  
 *** & *** & *** \\
\end{tabular}
\caption{Elaboration and innovation as predictors of beak shape distance from centroid.
The table shows parameter estimates of predictors from  13 different phylogenetic generalised least squares (PGLS) models with $\Delta$AIC showing relative model fit following removal of terms. Models 1-6 are multiple regressions including all variables and one or more interaction terms; models 7-9 are are multiple regressions including two or more variables and no interaction terms; models 10-13 are single predictor models. Values in the table are based on PGLS in which Pagel’s $\lambda$ was fixed at 1 (i.e. assuming a Brownian motion model of evolution). Fixing $\lambda$ is necessary to allow model comparison with AIC. We also fitted the same models fixing $\lambda$ to 0.7 (the lowest value of $\lambda$lambda from any individual model) and relative model fit and interpretation of parameters was unaffected. E = elaboration; I = innovation; AIC = Akaike Information Criterion; NS = not significant; - indicates that the term was not included within that model.}
\label{Tab:pgls}
\end{table}

@@

>>>>

Widely documented, megaevolutionary jumps in phenotypic diversity continue to perplex researchers because it remains unclear whether these dramatic changes can emerge from microevolutionary processes.
Here we tackle this question using new approaches for modeling multivariate traits %@L_multi_traits
 to evaluate the magnitude and distribution of elaboration and innovation in the evolution of bird beaks.
We find that elaboration, evolution along the major axis of phenotypic change, is common at both macro- and megaevolutionary scales whereas innovation, evolution away from the major axis of phenotypic change, is more prominent at megaevolutionary scales.
Indeed, the major axis of phenotypic change among species beak shapes at megaevolutionary scales is an emergent property of innovation across clades. Our analyses suggest that the reorientation of phenotypes via innovation is a ubiquitous route for divergence that can arise through gradual change alone, opening up new avenues for evolution to explore. %@L_reworded_abstract

@@

\begin{landscape}
\begin{table}[ht]
\centering
\begin{tabular}{rrllllllllrr}
  \hline
 Model & $E_{species}$ & $I_{species}$ & $E_{clade}$ & $I_{clade}$ & $E_{species}$ x $I_{clade}$ & $E_{species}$ x $E_{clade}$ & $I_{species}$ x $I_{clade}$ & $I_{species}$ x $E_{clade}$ & delta AIC & Adj. R2 \\ 
  \hline
1 & 0.04*** & -0.139*** & 0.002 & -0.002 & 0.064*** & 0.045*** & 0.341*** & 0.093*** & 0.00 & 0.92 \\ 
2 & 0.257*** & 0.169*** & 0.01 & 0.001 & 0.045*** &  & 0.312*** & 0.034*** & -691.48 & 0.92 \\ 
3 & 0.108*** & 0.236*** & 0.012 & -0.002 & 0.055*** & 0.034*** & 0.343*** &  & -284.74 & 0.92 \\ 
4 & 0.254*** & 0.104*** & 0.002 & 0.004 &  & 0.027*** & 0.311*** & 0.052*** & -877.85 & 0.92 \\ 
5 & 0.1*** & 0.571*** & -0.018* & 0.02*** & 0.054*** & 0.036*** &  & 0.096*** & -1498.36 & 0.91 \\ 
6 & 0.364*** & 0.973*** & -0.007 & 0.023*** &  &  &  &  & -2230.10 & 0.90 \\ 
7 & 0.364*** & 0.972*** &  &  &  &  &  &  & -2280.08 & 0.90 \\ 
8 &  &  & -0.025 & 0.011 &  &  &  &  & -22270.55 & -0.00 \\ 
9 & 0.286*** &  &  &  &  &  &  &  & -18986.63 & 0.32 \\ 
10 &  & 0.798*** &  &  &  &  &  &  & -17714.99 & 0.41 \\ 
11 &  &  & -0.008 &  &  &  &  &  & -22269.34 & -0.00 \\ 
12 &  &  &  & 0.002 &  &  &  &  & -22269.51 & -0.00 \\ 
   \hline
\end{tabular}
\caption{Elaboration and innovation as predictors of beak shape distance from centroid.
The table shows parameter estimates of predictors from  13 different phylogenetic generalised least squares (PGLS) models with $\Delta$AIC showing relative model fit following removal of terms. Models 1-6 are multiple regressions including all variables and one or more interaction terms; models 7-9 are are multiple regressions including two or more variables and no interaction terms; models 10-13 are single predictor models. Values in the table are based on PGLS in which Pagel’s $\lambda$ was fixed at 1 (i.e. assuming a Brownian motion model of evolution). Fixing $\lambda$ is necessary to allow model comparison with AIC. We also fitted the same models fixing $\lambda$ to 0.727 (the lowest value of $\lambda$lambda from any individual model) and relative model fit and interpretation of parameters was unaffected. E = elaboration; I = innovation; AIC = Akaike Information Criterion; NS = not significant; - indicates that the term was not included within that model.}
\label{Tab:pgls}
\end{table}
\end{landscape}

@@

\textbf{Author contributions.} G.H.T., A.P.B., N.C., and T. G. designed the study. J.A.B., C.R. C., E. C.H.,  and G.H.T. curated the data. T.G. analysed the data. All authors contributed to the writing of the manuscript.

\textbf{Competing interests.} We declare no competing interests.
